---
title: "Q2.2: How Age and diabetes affect the risk factors related to diabetes ?"
author: "Yaëlle Pihan, Marco Carega"
date: "13/04/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Table of contents
**
Setting up data and libraries  
1. How does age affect the risk factors associated with diabetes?  
2. Which of the risk factors are related with diabetes outcome?  
3. Two-Way Anova test considering "Blood Pressure".    
4. Analyze the distribution of Insulin variable.     
**

## Setting up data and libraries

```{r data, warning=FALSE, message=FALSE}
library(car)
library(carData)
library(sandwich)
library(lmtest)
library(RcmdrMisc)
library(agricolae)
diabetes=read.csv("diabetes.csv")

diabetes$AgeCat=factor(ifelse(diabetes$Age<=30,"young",ifelse(diabetes$Age<=50,"middle","old")))
diabetes$OutCat=factor(ifelse(diabetes$Outcome==1,"True","False"))

summary(diabetes)
```
## 1. How does age affect the risk factors associated with diabetes?
### a. Tests

```{r }
tests=function(factor){
    for (i in 1:7){
    print("----------------------------------------------------------")
    print(colnames(diabetes)[i])
    print("----------------------------------------------------------")
    print("----------------------Anova test--------------------------")
    model = aov(diabetes[,i]~factor,data=diabetes)
    print(summary(model))
    print("----------------------Assumption tests--------------------")
    print(shapiro.test(residuals(model))) #normality
    print(dwtest(model,alternative = "two.sided")) #independence
    print(bptest(model)) #homoscedasticity
    plot(density(diabetes[,i]), main = colnames(diabetes)[i]) #density plot to identify normality
    plot(model,2,sub = colnames(diabetes)[i]) #QQ plot to test normality
    plot(model,1,sub = colnames(diabetes)[i]) #Residuals vs fitted plot to test homoscedasticity
  }
}

tests(diabetes$Age)
```


### b. Analysis  
After analyzing the p_value of each Anova test, we conclude that the age of a patient affects significantly (with a p_value < 0.05) the risk factors below:

* Pregnancies
* Glucose
* Blood pressure  
* Skin thickness  

On the contrary (p_vale > 0.05), it does not impact:

* Insulin  
* BMI  
* Diabetes pedigree function

However, we need to be careful about some results because some Anova assumptions are more or less verified depending on the diabetes risk factor considered. 

Indeed, by opposition with the first part of Q2, the data we use is real and its groups do not necessary respect normality, independence or homoscedasticity. 

In our case, independence test is always verified: p_value > 0,05 and DW in [1.5;2.5], meaning the residuals are not linearly auto-correlated, so independence of the observations is verified. 

But normality (with Shapiro test) and homoscedasticity (with Durbin-Watson test) are never verified (p_value < 0.05). 

Therefore, we plot the residuals with a QQ plot showing us how much the residuals follow a linear regression. If we can identify easily a portion of straight line, we accept normality assumption. As for homoscedasticity, the residuals vs Fitted plot is supposed to show us equally spread residuals around the horizontal line. 

We choose to accept homoscedasticity assumptions except for Pregnancies.

However, we rejected normality assumption for Insulin, Skin thickness and Diabetes pedigree function. 

As a consequence, Anova results for Pregnancies, Insulin, Skin thickness and Diabetes pedigree function should be reconsidered and some transformations could be done on these dependent variables to try to make them fit Anova model. 

## 2. Which of the risk factors are related with diabetes outcome?
### a. Tests
```{r }
tests(diabetes$Outcome)
```

### b. Analysis
After analyzing the p_value of each Anova test, we conclude that the risk factors below are related significantly with diabetes outcome (p_value < 0,05):

* Pregnancies
* Glucose
* Skin thickness
* Insulin
* BMI
* Diabetes pedigree function

On the contrary, Blood pressure is not related with diabetes outcome (p_value > 0.05). 

Like before, we need to reconsider these results after analyzing if anova assumptions are verified.

About homoscedasticity assumptions, we choose to accept them all.

About normality, we choose to reject: Pregnancies, Insulin, Skin thickness and Diabetes pedigree function.  

### 3.1 Two-Way Anova test considering "Blood Pressure"  
Detail the results of Two-Way ANOVA considering “Blood Pressure” as dependent variable, and the age groups and the indicator of diabetes as independent variables.

### a. Anova Two way test  
Anova assumptions were tested before.
```{r }
model2 <- aov(diabetes$BloodPressure~diabetes$AgeCat+diabetes$OutCat, data = diabetes)
summary(model2)
TukeyHSD((model2))
plot(TukeyHSD((model2)))
```
### b. Analysis  
Thanks to Anova test results, we can conclude that the Age group category has a significant impact (*** and p_value<<0.05) on blood pressure whereas it is not the case for the indicator of diabetes (p_value = 0.931 > 0.05).

Moreover, Tukey Posthoc test gives the adjusted values for all possible pairs, and as they are all < 0.05, we can conclude that the difference of means come from all groups and explains anova results: Age groups have a significant impact on blood pressure values.  

## 3.2 Analyze the interaction term of two factors.  
### a. Interaction graph
```{r }
interaction.plot(diabetes$AgeCat,diabetes$OutCat, diabetes$BloodPressure, fun=mean, type="l", legend=TRUE)
model2 <- aov(diabetes$BloodPressure~diabetes$AgeCat*diabetes$OutCat, data = diabetes)
summary(model2)
```
### b. Analysis  
As we can see, the lines of diabetes indicator intercept, meaning there is likely an interaction effect between the age groups and diabetes indicator. However when we compute Anova with the interaction effect, the interaction is not significant enough (p_value = 0.281 > 0.05).   
## 4.1 Analyze the distribution of Insulin variable.  
```{r }
plot(density(diabetes$Insulin),main="Insulin")
#Boxplot(diabetes$Insulin~diabetes$AgeCat,data=data,id=FALSE)
#Boxplot(diabetes$Insulin~diabetes$OutCat,data=data,id=FALSE)
```
## 4.2 What would you recommend to fit an ANOVA model using Insulin levels as the dependent variable?
```{r}
diabetes$Insulin_log<- log(diabetes$Insulin)
plot(density(diabetes$Insulin_log),main="Insulin log")
# shapiro.test(diabetes$Insulin_log)
# diabetes["Insulin_log"][diabetes["Insulin_log"] == "-Inf"] <- -10000
# model2 <- aov(diabetes$Insulin_log~diabetes$AgeCat,diabetes)
# summary(model2)
```
In order to make Insulin fit an anova model, we could do a logarithmic transformation. 

